\LARGE \bf Unexpected Behaviors: Impact on Child-Robot Interaction

### 2015-02-27 paper.tex
Turned the paper into an IROS paper. Implemented a few of the HRI reviewer comments

\documentclass{sig-alternate}

@@

%
% --- Author Metadata here ---
\conferenceinfo{10th ACM/IEEE International Conference on Human-Robot Interaction}{2015 Portland, USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{\LARGE \bf
How Children Perceive and Interact with a Robot that Behaves Unexpectedly
}

%%% HRI 2015 -> double-blind review process

%\numberofauthors{1} 
%\author{
%\alignauthor
%Julia Fink, S\'{e}verin Lemaignan, Pierre Dillenbourg\\
%%\titlenote{Dr.~Trovato insisted his name be first.}
%       \affaddr{Computer-Human Interaction in Learning and Instruction Lab (CHILI)}\\
%       \affaddr{Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL)}\\
%       \affaddr{CH-1015 Lausanne}\\
%       \affaddr{Switzerland}\\
%       \email{firstname.lastname@epfl.ch}
%}
%
%\additionalauthors{Additional authors: 
%Francesco Mondada, LSRO, EPFL, francesco.mondada@epfl.ch 
%}
%

@@

Presented is a study on the impact of unexpected robot behaviours on the
perception of a robot by children and their subsequent engagement in a playful interaction.
We propose an original analysis methodology which blends behavioural cues and
reported phenomenological perceptions into a compound index.

@@

Besides, this study operationalizes for the first time the
\emph{anthropomorphism index} proposed by~\cite{fink2014dynamics}, which can be
considered as an effective tool to assess qualitative psycho-behavioural
episodes occurring during human-robot interaction.

@@

Similarly, looking at short-term engagement, \cite{short_no_2010} found that a
simple manipulation of the robot's behaviour can lead to greater engagement. The
authors let participants play several rounds of the rock-paper-scissors game
with the robot (the playfulness of the scenario seems important). When the robot
was cheating from time to time, participants tended to ascribe intention to the
robot what in turn led to greater engagement in how they were interacting with
the robot. The authors observe that \textit{``any deviation from expected
operation is sufficient to create a greater degree of engagement in the
interaction.''} Besides, they suggest that \textit{``many interactions can be
improved by the introduction of such simple behaviours, and that this should
be exploited by designers in HRI. Bringing human and robot together to
perform a simple, repetitive, familiar task and then having the robot behave
unexpectedly can increase engagement and mental state attribution without
complex behavioural or mechanical additions.''}  Along those lines, Fink
\etal\cite{fink2014dynamics} also suggest in their model of the dynamics of
anthropomorphism in HRI, that \emph{disruptive behaviours} may lead to increased
anthropomorphic projections and possibly engagement.

@@

This study also aims at assessing the effectiveness of the \emph{anthropomorphic
index} proposed in~\cite{fink2014dynamics} as a compound (behavioural
and phenomenological) metric for the attribution of human-like features to
a robot. We are especially interested in investigating how this indicator predicts
and/or reflects actual \emph{engagement} in the interaction.

@@

\begin{description}

    \item[Hypothesis 1:] Children show more engagement toward a robot that
        behaves unexpectedly from time to time compared to a robot that always
        behaves correctly.

    \item[Hypothesis 2:] Children perceive a robot that (tentatively) displays
        intention or cognitive abilities as more human-like than a robot that
        appears to have a system error, \ie the disobeying robot and the robot
        that makes a mistake will be more anthropomorphized than the robot that
        gets lost.

\end{description}

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bring things together and, relying on the concept of
anthropomorphism index~\cite{fink2014dynamics}, build a compound metric that
considers both children's perception and interaction aspects (qualitatively).
Previous work suggests that a social relation to a robot (we view
anthropomorphism as a specific type of social relation) reflects an increased
engagement and can be effective in sustaining interaction.  Consequently, we
formulate a third hypothesis:

\begin{description}

    \item[Hypothesis 3:] Anthropomorphic perception of the robot and the amount
    of engagement in the interaction positively correlate.

\end{description}

@@

In total, we obtained 2354 distinct actions which summed up to a total
annotation duration of 145 minutes. On average, one child accounted for 92
actions (SD=23).

@@

%We have measured engagement by combining \emph{interaction} data with the
%results of the semi-structured interviews.  Several of the aforementioned coded
%actions can reflect engagement: \textit{explore, misuse, gesture, touch, show,}
%and \textit{talk}. Also, \textit{look} at the experimenter can be considered as
%engagement in the interaction: We assume that children look at the experimenter
%when they are surprised and seek for help.  This behaviour reflects that they
%want to make sense of the robot's behaviour. Further, we can take into account
%how they refer to Ranger and describe the robot (as well as their experience) in
%the interviews.  Specifically, we used the constructs \textit{ascribe intention,
%ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe moral
%standing} (Table~\ref{tab:domino-questions}) to assess how far they engaged with
%the robot.
%
To account for the fact that anthropomorphism arises in an interaction, we bring
both children's perception of the robot (post-measurement) and their behaviour
toward it (in-the-moment measurement) together. We did so by building a
anthropomorphism index~\cite{fink2014dynamics}. We quantify the
index by attributing points for each anthropomorphic \emph{perception} of the
robot and for specific kinds of human-like \emph{behaviour}
toward the robot. The index included the following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable and the
weighting of each item can be discussed: we did not consistently assign 1 point
to each item, but assigned points between 0.5 and 2 points depending on our
perception of how a given item reflect on a higher level of anthropomorphic
perception of the robot (for instance, ascribing the ability to see and hear was
suggested by our study setup, and we cannot be sure that it really reflects
anthropomorphism). We believe however that the approach remains valid,
considering that the anthropomorphism index is primarily a \emph{comparison}
metric and not an absolute value.

@@

\paragraph{Emotions and Companionship}

We examined whether children attribute \textbf{feelings} to the robot by asking
them (once at the end of the experiment) if they thought that Ranger can feel
happy or sad sometimes. The majority of
children (21 of 26) gave a positive answer. Only 2 children who had interacted
with the \textit{disobeying} robot did not believe that Ranger has feelings and
another 3 children (2 \textit{disobey} and 1 \textit{lost}) did not reply at
all. Overall, data suggests that \textbf{children attribute emotional states}
to the robot. Asked for more details, several children answered that it is
through its colours and sounds that the robot shows a feeling. Most children said
that they could make the robot feel happy by playing with it and putting a
domino tile inside the box. This may also be a reflection of their own feelings,
projected on the robot. About half of the children (14 of 26) agreed that Ranger
can be their friend. We did not ask about what being a friend means to them but
as children generally liked playing with the robot, this may be linked to each
other. We can note that children \textbf{ascribe feelings to the robot and
partly accept that it can be a ``friend''}.

@@

We hypothesised that children who interact a lot and are probably more engaged
with the robot also perceive the robot as more human-like. Data suggests the
opposite, however. As shown in Figure~\ref{fig:domino-anthropo-interaction}, 
%(we even found a significant \textit{negative} correlation between the count of
%engagement actions (per group) and the qualitative anthropomorphism index --
%r(11)=-0.56, p=.05 -- but the sample size is too small to report it).  (b=-.05,
%t(11)=-2.22, p=.048).  The overall model with the count of engagement actions
%(per group) predicts a significant proportion of the qualitative
%anthropomorphism index (adjusted R$^2$=.25, F(1,11)=4.9, p=.05). This means
%that 
the more a group showed engagement in the interaction, the less they
anthropomorphized the robot. This is a key result, which was against our
initial assumption. A possible interpretation is that children who interact more
with the robot understand better how it works, they are more familiar with it,
and as such the robot appears less ``mystical'' to them, so there is not a big
need to anthropomorphize it. On the contrary, the cluster of groups that do not
interact much but anthropomorphize the robot more, is quite homogeneous, and may
reflect a certain fear of interacting with a robot that would look to them to
human-like.
%The negative correlation between the number of engagement actions and the
%anthropomorphism index suggests that children who interact more with the robot
%tend to anthropomorphize it less. This implies that anthropomorphism could fade
%out after some time (or rather after some interaction).
This raises the question how far anthropomorphism (as a special kind of social
engagement) really helps in sustaining interaction. This is a critical point
because most of the short-term investigations suggest that anthropomorphic
design and human social cues emitted by a robot foster engagement and
acceptance. What if this is not true for continued interaction, and thus for the
long-term? We need to remain modest here: while we found a significant
\emph{negative} correlation between engagement and anthropomorphism in the data
pictured on Figure~\ref{fig:domino-anthropo-interaction} (r(11)=-0.56, p=.05),
we have to be careful about our interpretation, due to the small sample size (13
pairs), and we suggest to investigate the aspect further in future research.

@@

As hypothesised, we found that in a playful scenario where 4-5 year old children
play domino together with a robot, the robot seems to be more engaging when it
shows some misbehaviour compared to when it always behaves as expected
(notwithstanding the likely impact of a novelty effect).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
%
% double blind review
%This research was supported by the Swiss National Science Foundation through the
%National Centre of Competence in Research Robotics.
We would like to thank all the families who participated to the study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrv}
\bibliography{domino}

\balancecolumns

>>>>

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\overrideIEEEmargins

@@

\title{\LARGE \bf
Unexpected Behaviors: Impact on Child-Robot Interaction
}

\author{
    S\'{e}verin Lemaignan$^{1,2}$, Julia Fink$^{1}$, Francesco Mondada$^{2}$, Pierre
    Dillenbourg$^{1}$\\
$^{1}$Computer-Human Interaction in Learning and Instruction Lab (CHILI) \\
$^{2}$Laboratoire de Systèmes Robotiques (LSRO) \\
École Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\
CH-1015 Lausanne, Switzerland \\
{\tt\small firstname.last@epfl.ch}
}

@@

We present a study on the impact of unexpected robot behaviours on the
perception of a robot by children and their subsequent engagement in a playful
interaction based on a novel "domino" task.
We propose an original analysis methodology which blends behavioural cues and
reported phenomenological perceptions into a compound index.

@@

@@

Similarly, looking at short-term engagement, \cite{short_no_2010} found that a
simple manipulation of the robot's behaviour can lead to greater engagement. The
authors let participants play several rounds of the rock-paper-scissors game
with the robot (the playfulness of the scenario seems important). When the robot
was cheating from time to time, participants tended to ascribe intention to the
robot what in turn led to greater engagement in how they were interacting with
the robot. The authors observe that \textit{``any deviation from expected
operation is sufficient to create a greater degree of engagement in the
interaction.''} Besides, they suggest that \textit{``many interactions can be
improved by the introduction of such simple behaviours, and that this should
be exploited by designers in HRI. Bringing human and robot together to
perform a simple, repetitive, familiar task and then having the robot behave
unexpectedly can increase engagement and mental state attribution without
complex behavioural or mechanical additions.''}  Along those lines, we also
suggested in our model of the dynamics of anthropomorphism in
HRI~\cite{lemaignan2014dynamics} that \emph{disruptive behaviours} may lead to
increased anthropomorphic projections and possibly increased engagement.

@@

The main outcomes of this research, detailed hereafter, can be summarized as:

\begin{enumerate}
    \item an actionable approach (introduction of mis-behaviours) to support
        child-robot engagement,
    \item a mixed technique, blending behavioural cues and reported
        phenomenological perceptions, to assess the robot perception in terms of
        both engagement and human-likeliness,

    \item a new experimental protocol that compares three types of
        mis-behaviours, with different cognitive correlates,

    \item and a first experimental cue that anthropomorphic perceptions do not
        necessarily correlate with actual engagement.
\end{enumerate}

@@

    \item {\bf Hypothesis 1}: Children show more engagement toward a robot that
        behaves unexpectedly from time to time compared to a robot that always
        behaves correctly.

    \item {\bf Hypothesis 2}: Children perceive a robot that (tentatively) displays
        intention or cognitive abilities as more human-like than a robot that
        appears to have a system error, \ie the disobeying robot and the robot
        that makes a mistake will be more anthropomorphized than the robot that
        gets lost.

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bring things together and propose a compound metric that
considers both children's perception and interaction aspects (qualitatively).
Previous work suggests that a social relation to a robot (we view
anthropomorphism as a specific type of social relation) reflects an increased
engagement and can be effective in sustaining interaction.  Consequently, we
formulate a third hypothesis:

\item {\bf Hypothesis 3}: Anthropomorphic perception of the robot and the amount
    of engagement in the interaction positively correlate.

@@

In total, we obtained 2354 distinct actions. On average, interactions lasted
13min 43sec, and one child accounted for 90.53 actions (SD=23).

@@

%We have measured engagement by combining \emph{interaction} data with the
%results of the semi-structured interviews.  Several of the aforementioned coded
%actions can reflect engagement: \textit{explore, misuse, gesture, touch, show,}
%and \textit{talk}. Also, \textit{look} at the experimenter can be considered as
%engagement in the interaction: We assume that children look at the experimenter
%when they are surprised and seek for help.  This behaviour reflects that they
%want to make sense of the robot's behaviour. Further, we can take into account
%how they refer to Ranger and describe the robot (as well as their experience) in
%the interviews.  Specifically, we used the constructs \textit{ascribe intention,
%ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe moral
%standing} (Table~\ref{tab:domino-questions}) to assess how far they engaged with
%the robot.
%
To account for the fact that anthropomorphism arises in an interaction, we bring
both children's perception of the robot (post-measurement) and their behaviour
toward it (in-the-moment measurement) together. We did so by building a
so-called \emph{anthropomorphism index}. We quantify the
index by attributing points for each anthropomorphic \emph{perception} of the
robot and for specific kinds of human-like \emph{behaviour}
toward the robot. The index included the following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable and the
weighting of each item can be discussed: we did not consistently assign 1 point
to each item, but assigned points between 0.5 and 2 points depending on our
perception of how a given item reflect on a higher level of anthropomorphic
perception of the robot (for instance, ascribing the ability to see and hear was
suggested by our study setup, and we cannot be sure that it really reflects
anthropomorphism). Note that these issues are mitigated by using the
anthropomorphism index as a \emph{relative} metric (comparison between
conditions) and not an absolute value.

@@

%\paragraph{Emotions and Companionship}
%
%We examined whether children attribute \textbf{feelings} to the robot by asking
%them (once at the end of the experiment) if they thought that Ranger can feel
%happy or sad sometimes. The majority of
%children (21 of 26) gave a positive answer. Only 2 children who had interacted
%with the \textit{disobeying} robot did not believe that Ranger has feelings and
%another 3 children (2 \textit{disobey} and 1 \textit{lost}) did not reply at
%all. Overall, data suggests that \textbf{children attribute emotional states}
%to the robot. Asked for more details, several children answered that it is
%through its colours and sounds that the robot shows a feeling. Most children said
%that they could make the robot feel happy by playing with it and putting a
%domino tile inside the box. This may also be a reflection of their own feelings,
%projected on the robot. About half of the children (14 of 26) agreed that Ranger
%can be their friend. We did not ask about what being a friend means to them but
%as children generally liked playing with the robot, this may be linked to each
%other. We can note that children \textbf{ascribe feelings to the robot and
%partly accept that it can be a ``friend''}.

@@

We hypothesised that children who interact a lot and are probably more engaged
with the robot also perceive the robot as more human-like. Data suggests the
opposite, however. As shown in Figure~\ref{fig:domino-anthropo-interaction}, 
%(we even found a significant \textit{negative} correlation between the count of
%engagement actions (per group) and the qualitative anthropomorphism index --
%r(11)=-0.56, p=.05 -- but the sample size is too small to report it).  (b=-.05,
%t(11)=-2.22, p=.048).  The overall model with the count of engagement actions
%(per group) predicts a significant proportion of the qualitative
%anthropomorphism index (adjusted R$^2$=.25, F(1,11)=4.9, p=.05). This means
%that 
the more a group showed engagement in the interaction, the less they
anthropomorphized the robot. This is a key result, which was against our
initial assumption. A possible interpretation is that children who interact more
with the robot understand better how it works, they are more familiar with it,
and as such the robot appears less ``mystical'' to them, so there is not a big
need to anthropomorphize it. On the contrary, the cluster of groups that do not
interact much but anthropomorphize the robot more, is quite homogeneous, and may
reflect a certain fear of interacting with a robot that would look to them too
human-like.
%The negative correlation between the number of engagement actions and the
%anthropomorphism index suggests that children who interact more with the robot
%tend to anthropomorphize it less. This implies that anthropomorphism could fade
%out after some time (or rather after some interaction).
This raises the question how far anthropomorphism (as a special kind of social
engagement) really helps in sustaining interaction. This is a critical point
because most of the short-term investigations suggest that anthropomorphic
design and human social cues emitted by a robot foster engagement and
acceptance. What if this is not true for continued interaction, and thus for the
long-term? We need to remain modest here: while we found a significant
\emph{negative} correlation between engagement and anthropomorphism in the data
pictured on Figure~\ref{fig:domino-anthropo-interaction} (r(11)=-0.56, p=.05),
we have to be careful about our interpretation, due to the small sample size (13
pairs), and we suggest to investigate the aspect further in future research.

@@

As hypothesised, we found that in a playful scenario where 4-5 year old children
play domino together with a robot, the robot seems to be more engaging when it
shows some misbehaviour compared to when it always behaves as expected
(notwithstanding the impact of a novelty effect).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

We would like to thank all the families who participated to the study.
This research was supported by the Swiss National Science Foundation through the
National Centre of Competence in Research Robotics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtran}
\bibliography{domino}
